---
title: "R `dplyr` to Python `pandas`"
author: "Lukas Warode"
subtitle: A Translation using Gapminder Data
output:
  html_document:
    toc: yes
    toc_float: yes
    toc_depth: 2
---

The aim of this document is to compare or rather translate the syntax of R's `dplyr` to Python's `pandas`.

For this objective, the Gapminder data source is used, backed by the `gapminder` package, while the original content is inspired by [this source](https://gist.github.com/conormm/fd8b1980c28dd21cfaf6975c86c74d07#file-r-to-python-data-wrangling-basics-md).

```{r rmd setup, include=FALSE}
knitr::opts_chunk$set(error = FALSE,
                      message = FALSE,
                      warning = FALSE)
```

```{r packages}
library(tidyverse)
library(gapminder)
library(reticulate)
```

## Python configuration

```{r prerequisites, message=FALSE, results = "hide"}
use_python("/usr/local/bin/python")
Sys.setenv("RETICULATE_PYTHON" = "/Users/elweeh/Library/r-miniconda/envs/r-reticulate/bin/python3.7")
py_config()
```

```{python, message=FALSE}
import pandas as pd
import numpy as np
```

## Gapminder data

### R

```{r df}
df <- gapminder

if (! "gapminder.csv" %in% list.files()) {
  
  write_csv(df, "gapminder.csv")
  
}

# take a look at gapminder data
DT::datatable(df)
```

### Python

```{python, message=FALSE}
df = pd.read_csv("gapminder.csv")
```


## `filter`

### Example 1

#### R

```{r}
df %>% filter(lifeExp > 60 & lifeExp < 80) # filter(df, lifeExp > 60 & lifeExp < 80) 
```

#### Python

```{python}
df[(df["lifeExp"] > 60) & (df["lifeExp"] < 80)]
```

### Example 2

#### R

```{r}
df %>% filter(country == "Germany") # filter(df, country == "Germany") 
```

#### Python

```{python}
df[df["country"] == "Germany"]
```

### Example 3

#### R

```{r}
df %>% group_by(continent, year) %>% filter(mean(gdpPercap) > 10000)
```

#### Python

```{python}
df.groupby(['continent', 'year']).filter(lambda x: np.mean(x['gdpPercap']) > 10000)
```

## `select`

### Example 1

#### R

```{r}
df %>% select(country, pop) # select(df, country, pop) 
```

#### Python

```{python}
df[["country", "pop"]]
```

### Example 2

#### R

```{r}
df %>% select(- continent) # select(df, - continent) 
```

#### Python

```{python}
df.drop("continent", axis = 1)
```

## `arrange`

### Example 1

#### R

```{r}
df %>% arrange(lifeExp) # arrange(df, lifeExp)
```

#### Python

```{python}
df.sort_values("lifeExp")
```

### Example 2

#### R

```{r}
df %>% arrange(desc(lifeExp)) # arrange(df, desc(lifeExp))
```

#### Python

```{python}
df.sort_values("lifeExp", ascending = False)
```

## `distinct`

### Example 1

#### R

```{r}
df %>% distinct() # distinct(df)
```

#### Python

```{python}
df.drop_duplicates()
```

### Example 2

#### R 

```{r}
df %>% distinct(continent) # distinct(df, continent)
```

#### Python

```{python}
df.drop_duplicates(subset = "continent")
```

## `sample`

### Example 1

#### R 

```{r}
df %>% sample_n(100) # sample_n(df, 100)
```

#### Python

```{python}
df.sample(100)
```

### Example 2

#### R 

```{r}
df %>% sample_frac(0.5) # sample_n(df, 0.5)
```

#### Python

```{python}
df.sample(frac = 0.5)
```

## `group_by`

### Example 1

#### R

```{r}
df %>% group_by(continent) # group_by(df, continent)
```

#### Python

```{python}
df.groupby("continent")
```

### Example 2

#### R

```{r}
df %>% ungroup() # ungroup(df)
```

#### Python

```{python}
df.reset_index()
```

## `group_by` and `summarise`

### Example 1

#### R

```{r}
df %>% 
  group_by(year) %>% 
  summarise(mean_lifeExp = mean(lifeExp))
```

#### Python

```{python}
df.groupby("year")["lifeExp"].mean().reset_index(name = "mean_lifeExp") # df.groupby("year")["lifeExp"].agg(mean_lifeExp = ("lifeExp", "mean"))
```

### Example 2

#### R

```{r}
df %>% 
  group_by(continent, year) %>% 
  summarise(mean_gdpPercap = mean(gdpPercap), 
            sum_pop = sum(gdpPercap), 
            country_n = n())
```

#### Python

```{python}
df.groupby(["continent", "year"])["gdpPercap"].agg(["mean", "sum", "count"])
```

### Example 3

#### R

```{r}
df %>% 
  group_by(continent, year) %>% 
  summarise(mean_gdpPercap = mean(gdpPercap), 
            sum_pop = sum(pop), 
            country = first(country))
```

#### Python

```{python}
df.groupby(["continent", "year"]).agg(mean_gdpPercap = ("gdpPercap", "mean"), sum_pop = ("pop", "sum"), country = ("country", "first"))  
```

## `mutate`

### Example 1

### Example 2





